---
title: "Generating a DEM for a polygon"
author: "GG"
date: "6/26/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


I have found the elevatr package really useful for downloading digital elevation models (DEMs) for areas that you are interested in. In this short tutorial, I will demonstrate how to download a DEM using the elevatr package and then clip it to a polygon of interest, in this case a watershed. [_Click here_](https://rdrr.io/cran/elevatr/) for more information on the elevatr package. 

To get started make sure you have a spatial polygon that represents the area you are interested in. This can be a .shp file from ESRI ArcGIS or a vector file from other programs. For this tutorial it just needs to be able to be read by the [_sf package_](https://r-spatial.github.io/sf/). I am going to use a watershed in Iowa.

Load packages
```{r, echo = T, eval = T, cache = F, message = F, warning = F}
library(sf)
library(mapview)
library(elevatr)
library(raster)
```


Read in the watershed outline
```{r, echo = T, eval = T, cache = F, message = F, warning = F}
setwd("~/Google Drive File Stream/My Drive/UCSC Google Drive/Projects/Extra/Website/Soil_Example/")
vm <- st_read("vanmeter.shp")
```

View the watershed
```{r, echo = T, eval = T, cache = F, message = F, warning = F}
mapview(vm)
```

Download the DEM, the z parameter controls the resolution of the DEM. For the size of this watershed ~9,000 km2 we'll use z = 10, which will give us a roughly 50m resolution
```{r, echo = T, eval = T, cache = F, message = F, warning = F}
vm_dem <- get_elev_raster(vm,z=10)
```

View the watershed overlaid on the downloaded DEM
```{r, echo = T, eval = T, cache = F, message = F, warning = F}
mapview(vm_dem)+
  mapview(vm, col.regions = 'transparent', alpha.regions = 0, lwd = 3)
```

Clip the DEM to the outline of the watershed the mask function from the raster package
```{r, echo = T, eval = T, cache = F, message = F, warning = F}
vm_clipped <- mask(vm_dem, vm)
```

Now view the final results
```{r, echo = T, eval = T, cache = F, message = F, warning = F}
mapview(vm_clipped, na.color = 'transparent')+
  mapview(vm, col.regions = 'transparent', alpha.regions = 0, lwd = 3)
```

